---
title: 'NYC Squirrel Census - #TidyTuesday'
author: "Arthur Cheib"
date: "14/11/2019"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(statebins)
library(cowplot)
library(lubridate)
library(showtext)
library(hrbrthemes)
library(RColorBrewer)
library(ggthemes)
library(ggrepel)

font_add_google("Staatliches")
font_add_google("Oswald")
font_add_google("Montserrat", regular.wt = 300, bold.wt = 500)
```

## Get the data

```{r}
tb_lifts <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-08/ipf_lifts.csv") %>% 
  mutate(year2 = year(date))

```

## Question 01:

1) What was the evolution of the max lifted weigth over the years on bench and deadlift categories (divided by sex and desconsidering the weight class division)? 

```{r}
final_plot1 <- tb_lifts %>%
  group_by(year(date), sex) %>% 
  summarize(MAX_BENCH = max(best3bench_kg, na.rm = T),
            MAX_DEADLIFT = max(best3deadlift_kg, na.rm = T)) %>% 
  ungroup() %>% 
  rename("year" = `year(date)`,
         "Sex" = "sex") %>%
  arrange(Sex, year) %>%
  gather(key = class_type, value = weight_lifed, -year, -Sex) %>% 
  mutate(class_type = fct_recode(class_type, "BENCH CATEGORY" = "MAX_BENCH")) %>% 
  mutate(class_type = fct_recode(class_type, "DEADLIFT CATEGORY" = "MAX_DEADLIFT")) %>%
  ggplot(aes(x = year, y = weight_lifed, group = Sex, color = Sex)) +
  geom_line(size = 1.25) +
  facet_wrap(~class_type) +
  theme_wsj() +
  geom_point(shape = 21, color = "white", fill = "#69b3a2", size = 3) +
  labs(title = "How stronger have we become?",
       subtitle = "An evolution of the maximum weight lifted over the years - kg") +
  ylim(100, 450) +
  theme(legend.position = "top",
        plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 16),
        plot.caption = element_text(size = 12)) 


final_plot1

```

## Question 02:

1) Which weigth category presented the greatest prop evolution in the max lifted weigth over the years on bench and deadlift? 

```{r}
### Discovering and considering only the weigth classes that appeared over the last five decades of competition:
all_five_year_appearance <- tb_lifts %>%
  mutate(half_decade = year(date) - year(date) %% 5) %>%
  count(half_decade, weight_class_kg) %>%
  count(weight_class_kg, sort = T) %>% 
  filter(n > 8) %>% 
  pull(weight_class_kg)

### Treating data
tb_quest2 <- tb_lifts %>%
  filter(weight_class_kg %in% all_five_year_appearance) %>%
  mutate(half_decade = year(date) - year(date) %% 5) %>%
  group_by(half_decade, sex, weight_class_kg) %>% 
  summarize(MAX_BENCH = max(best3bench_kg, na.rm = T),
            MAX_DEADLIFT = max(best3deadlift_kg, na.rm = T)) %>% 
  ungroup() %>% 
  arrange(half_decade, sex) %>%
  mutate(PROP_MAX_BENCH = round(as.numeric(MAX_BENCH)/as.numeric(weight_class_kg), digits = 2),
         PROP_MAX_DEADLIFT = round(as.numeric(MAX_DEADLIFT)/as.numeric(weight_class_kg), digits = 2)) %>% 
  select(-MAX_BENCH, -MAX_DEADLIFT) %>% 
  gather(key = exercise_type, value = prop_weight_lifed, -half_decade, -sex, -weight_class_kg)


### Creating a mean weight lifted for faceting
mean_weight <- tb_quest2 %>%
  filter(exercise_type == "PROP_MAX_BENCH") %>% 
  group_by(weight_class_kg) %>% 
  summarize(prop_weight_lifed = mean(prop_weight_lifed)) %>% 
  ungroup()

mean_weight2 <- tb_quest2 %>%
  filter(exercise_type == "PROP_MAX_BENCH") %>% 
  group_by(weight_class_kg) %>% 
  summarize(prop_weight_lifed = mean(prop_weight_lifed)) %>% 
  ungroup() %>% 
  as_tibble() %>% 
  mutate(sex = "F") %>% 
  bind_rows(tb_quest2 %>%
  filter(exercise_type == "PROP_MAX_BENCH") %>% 
  group_by(weight_class_kg) %>% 
  summarize(prop_weight_lifed = mean(prop_weight_lifed)) %>% 
  ungroup() %>% 
  as_tibble() %>% 
  mutate(sex = "M"))

### Plotting Bench Max
tb_quest2 %>%
  filter(half_decade != 2015) %>% 
  filter(exercise_type == "PROP_MAX_BENCH") %>%
  mutate(weight_class_kg_f = factor(x = weight_class_kg,
                                  levels = c("52", "56", "60", "67.5", "75", "82.5", "90", "100", "110"))) %>% 
  arrange(half_decade, weight_class_kg) %>% 
  ggplot(aes(x = half_decade, y = prop_weight_lifed, color = sex)) +
  geom_line(size = 1.25) +
  facet_wrap(~weight_class_kg, scales = "free_y") +
  theme_ipsum() +
  geom_point(shape = 21, color = "white", fill = "#69b3a2", size = 3) +
  ggtitle(label = "Evolution of Weigth Records across the Years",
          subtitle = "by weight categories") +
  ylim(1, 5) +
  scale_x_continuous(name = "# Each 5 Years", breaks = seq(1970, 2015, by = 5)) +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(aes(yintercept = prop_weight_lifed), mean_weight,
             colour = "orange",
             size = 0.8,
             linetype = "dashed") +
  geom_text(aes(x = 1980, y = 2, label = str_c(round(prop_weight_lifed, digits = 2), "x Weight's Class \n(Mean)")),
                data = mean_weight2,
                vjust = -1, colour = "orange")


```

## Question 03:

1) Is there any championship in which we can expect to watch world records on Bench type competition?

```{r}

tb_lifts %>%
  mutate(half_decade = year(date) - year(date) %% 5) %>%
  group_by(half_decade, sex) %>% 
  summarize(MAX_BENCH = max(best3bench_kg, na.rm = T)) %>% 
  ungroup() %>% 
  arrange(sex, half_decade) %>%
  left_join(tb_lifts %>% 
              select(sex, best3bench_kg, year2, meet_name),
            by = c("sex", "MAX_BENCH" = "best3bench_kg")) %>% 
  arrange(year2) %>% 
  group_by(half_decade, sex) %>% 
  slice(1) %>% 
  ungroup() %>% 
  select(-year2) %>% 
  ggplot(aes(x = half_decade, y = MAX_BENCH, label = meet_name)) +
  geom_point(size = 3) +
  geom_label_repel(aes(fill = factor(sex)), colour = "#D1E5F0", fontface = "bold") +
  theme_modern_rc() +
  labs(title = "In which competitions world's records are broken?",
       subtitle = "... in world competitions (mostly)",
       x = "",
       y = "Lifted Weight (Kg)",
       caption = "#TidyTuesday post by @Arthur Cheib  |  Data from Open Powerlifting.") +
  theme(legend.position = "none",
        plot.title = element_text(size = 20),
        plot.subtitle = element_text(size = 16),
        plot.caption = element_text(size = 12)) +
  ylim(100, 450)

```
